---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hoopTotals

<!-- badges: start -->
<!-- badges: end -->

There is a lot of research available for win probability models in various sports.
However, there is much less research in predicting the number of points scored in a particular game.

This package implements the method described in the reference below, but 
for NCAA Men's Basketball, rather than the NBA.

Reference:
Song, K., Y. Gao, and J. Shi. 2020. 'Making Real-Time Predictions for NBA Basketball Games 
by Combining the Historical Data and Bookmaker’s Betting Line.' Physica A 547: 1–8


## Installation

You can install the development version of hoopTotals from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("statistnick-tamu/hoopTotals", build_vignettes = TRUE)
```


## Functions


total: calculates expected points scored for a game

live: adjusts expected points based on what has happened in the game

train: calculates parameters to pass into total & live for specific teams

plotTotal: creates a hypothetical scoring flow plot


## Example

Here are a few quick examples:


```{r example}
library(hoopTotals)
total(3.5,5)
```


```{r}
live(3.5, 5, 20, 76, 152)
```


```{r plot}
plotTotal(3.5, 5, 140, 144.5)
```
