---
title: "my-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hoopTotals)
```

# Using hoopTotals

There is a lot of research available for win probability models in various sports.
However, there is much less research in predicting the number of points scored in a particular game.

This package implements the method described in 'Making real-time predictions for NBA basketball games by combining the historical data and bookmaker’s betting line' for NCAA Men's Basketball, rather than the NBA.

# total

The 'total' function implements the method and returns expected points.
It needs an estimate of the average and variance of points scored per minute
for the two teams you are interested in estimating a score for.

3.5 and 5 are used by default as they are the averages from the last two seasons of NCAA Men's Basketball

```{r total1}
total(3.5, 5)
```

We would expect two teams that score around 3.5 points per minute (combined) to score 140 points in a game.

Now, let's try adding in some additional information.

# improving total

```{r total2}
total(3.5, 5, time = 20, current = 80, threshold = 160.5)
```

We have provided additional information saying that the game is 20 minutes in (halfway through), there 
have been 80 points scored so far, and we want to know the probability we see more than 160 points.

The function still outputs an expected total, but that has now increased to about 150. Also output is the
probability that more than 160 points are scored and what the model would consider to be the betting line.

The authors also included an adjusted method which takes into account the line set by the market
prior to the game.

```{r total3}
total(3.5, 5, time = 20, current = 80, threshold = 160.5, line = 152.5)
```

If we provide that additional information saying that the market expected 152.5 points to be scored,
you can see the expected points are drawn upwards and the probability of hitting the threshold increases.

That's all well and good, but we're still making a general assumption around how much a team scores.
We can do better than that.

# train

In the 'train' function, you can provide play-by-play data frames for two teams
and it will calculate the first two parameters to pass back in to the 'total' function.

I've loaded up data from Texas A&M and Iowa State's Men's Basketball teams here.

```{r data}
head(home.df)
head(away.df)
train(home.df, away.df)
```

Let's pass those into total and see how it changes.

```{r total4}
total(4.5375, 4.8663)
```

We expect a much higher scoring game now. 

The two teams played on 11/26/23 and had a much lower expected total.

```{r total5}
total(4.5375, 4.8663, time=40, current=0, threshold=135.5, line=135.5)
```

If we feed in the pre-game line, we see a drastic decrease in expected points.

At halftime, there were 70 points scored, let's see how that changes the expectation.

```{r total6}
total(4.5375, 4.8663, time=20, current=70, threshold=135.5, line=135.5)
```

Now, we're expecting 138 points and a 62% chance of clearing the pre-game line.

# plotTotal

Lastly, you can plot a hypothetical game flow using the plotTotal function. 

You'll have to provide some similar information as the other functions:

```{r plot}
plotTotal(ppm=4.5375, var=4.8663, threshold=135.5, line=135.5)
```
