---
title: "hoopTotal-how-to"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hoopTotal-how-to}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hoopTotals)
```


## total

The 'total' function implements the method and returns expected points.
It needs an estimate of the average and variance of points scored per minute
for the two teams you are interested in estimating a score for.

3.5 and 5 are used by default as they are the averages from the last two seasons of NCAA Men's Basketball.

```{r total1}
total(3.5, 5)
```

We would expect two teams that score around 3.5 points per minute (combined) to score 140 points in a game.

Now, let's try adding in some additional information.


## improving total

```{r total2}
total(3.5, 5, line = 150)
```

The authors also included an adjusted method which takes into account the line set by the market
prior to the game. Here, we have provided extra information by saying the market expects
150 points to be scored in this game.

As you can see, the expected points moves aggressively towards that.


## live

```{r total3}
live(3.5, 5, time = 20, current = 70, threshold = 140, line = 150)
```

The interesting part of the method is the live component. Here, we have added a few
more pieces of information: how much time has passed, how many points are currently scored,
and a user defined threshold (this could be a betting line, or a prior expected score).

This function still returns the expected points but now calculates a probability that
the threshold is exceeded and also converts that probability into a betting line (price).

This is great, but we're still making a general assumption around how much certain teams score.
We can do better than that.


## train

In the 'train' function, you can provide play-by-play data in the form of a data frame for two teams
and it will calculate the first two parameters to pass back in to the other functions. See the 'train' help text for specifications around how to input the data.


Included with the package are data from Texas A&M and Iowa State's Men's Basketball teams from the 2022-23 season and
partial from the current 2023-24 season.

```{r data}
head(home.df)
head(away.df)
train(home.df, away.df)
```


Let's pass those into total and see how it changes.

```{r total4}
total(4.5375, 4.8663)
```

We expect a much higher scoring game now. 

However, the two teams played on 11/26/23 and had a much lower expected total.
If we feed in the pre-game line, we see a drastic decrease in expected points.

```{r total5}
total(4.5375, 4.8663, line=135.5)
```


At halftime, there were 70 points scored, let's see how that changes the expected points.

```{r total6}
live(4.5375, 4.8663, time=20, current=70, threshold=135.5, line=135.5)
```

Now, we're expecting 137 points and a 62% chance of clearing the pre-game line.


## plotTotal

Lastly, you can plot a hypothetical game flow using the plotTotal function. 
You'll have to provide similar information as the other functions:

```{r plot}
plotTotal(ppm=4.5375, var=4.8663, threshold=135.5, line=135.5)
```
